---
title: "Descriptive Statistics"
author: "Dean Hawk"
date: "2024-07-23"
output: html_document
---

## Table of Contents:Several Methods for Descriptive Statistics
1.Custom Functions
• Custom Descriptive Statistics Function:Create a function to calculate basic descriptive statistics such as mean,median,standard deviation,etc.

2.Base R Functions
• `summary()`:Use the base R function`summary()`to obtain descriptive statistics for the data.

3.Functions from R Packages
• `psych::describe()`:Use the`describe()`function from the`psych`package to obtain more detailed descriptive statistics.
• `pastecs::stat.desc()`:Use the`stat.desc()`function from the`pastecs`package to obtain descriptive statistics,including skewness and kurtosis.
• `skimr::skim()`:Use the`skim()`function from the`skimr`package to obtain comprehensive descriptive statistics.
• `bruceR::Describe()`:Use the`Describe()`function from the`bruceR`package to obtain descriptive statistics.
• `Hmisc::describe()`:Use the`describe()`function from the`Hmisc`package to obtain descriptive statistics.

```{r}
# Clear the environment
rm(list = ls()) 

# Load necessary libraries：commonly used R packages for descriptive statistics
# install.packages("pacman")
pacman::p_load(
  tidyverse,
  Hmisc,
  psych,
  skimr,
  pastecs,
  bruceR
)

# Use str() to display the structure of the mtcars dataset
str(mtcars)

# Use glimpse() to display the structure of the mtcars dataset
glimpse(mtcars)
```


## 1.Custom Function

```{r Custom Function}
# Create a function for descriptive statistics
my_describe <- function(x){
  options(digits = 3)  # Set the number of digits to display
  N = length(x);  # Number of observations
  Min = min(x, na.rm = TRUE);  # Minimum value
  Q1 = quantile(x, probs = 0.25, na.rm = TRUE);  # First quartile (25th percentile)
  Median = median(x, na.rm = TRUE);  # Median (50th percentile)
  Q3 = quantile(x, probs = 0.75, na.rm = TRUE);  # Third quartile (75th percentile)
  Max = max(x, na.rm = TRUE);  # Maximum value
  Mean = mean(x, na.rm = TRUE);  # Mean
  Var = var(x, na.rm = TRUE);  # Variance
  Sd = sd(x, na.rm = TRUE);  # Standard deviation
  Range = abs(diff(range(x)));  # Range (Max - Min)
  # Return the results as a data frame
  return(data.frame(N = N, Min = Min, Q1 = Q1, Median = Median, Q3 = Q3, Max = Max, Mean = Mean, Var = Var, Sd = Sd, Range = Range))
}

# Apply the function to each column of the mtcars dataset
sapply(mtcars, my_describe)
```

## 2. Base R ：summary()   Function

The summary() function provides the minimum, first quartile, median, third quartile, and maximum for numeric variables.

```{r summary()}
# Display the first few rows of the mtcars dataset
head(mtcars)

# Select specific columns for analysis
data <- mtcars

# Use the summary() function
summary(data)
```

## 3.psych Package describe() Function

```{r psych::describe()}
# pacman::p_load(psych)
psych::describe(data)
```

## 4.pastecs包中stat.desc()函数

```{r pastecs::stat.desc()}
# Install the pastecs package if not already installed
# install.packages("pastecs")
pacman::p_load(pastecs)

# Set the number of digits to display
options(digits=3) 

# Use the stat.desc() function with norm = TRUE
stat.desc(data, norm = TRUE)
```

## 5.skimr::skim()

```{r}
# pacman::p_load(skimr)
skimr::skim(data)
```

## 6.bruceR::Describe()

```{r}
# pacman::p_load(bruceR)
bruceR::Describe(mtcars)
```

## 7.Hmisc::describe()

```{r}
# pacman::p_load(Hmisc)
Hmisc::describe(mtcars)
```




















