---
title: "Exploratory Data Analysis(EDA)"
author: "Dean Hawk"
date: "2024-07-23"
output: html_document
---

## Table of Contents：Exploratory Data Analysis(EDA)
	•	Graphics Package
	•	DataExplorer Package
	•	GGally Package: ggpairs()

## Graphics Package

1.Bar Plot(`barplot()`)
2.Box Plot(`boxplot()`)
3.Histogram(`hist()`)
4.Scatter Plot Matrix(`pairs()`)

```{r}
# Load necessary packages
rm(list = ls())  # Clear the environment
# install.packages("pacman")
pacman::p_load(
  tidyverse,
  graphics
)

# Data import
data <- mtcars

# 1. Bar Plot (barplot())
# Create a bar plot for the number of cylinders in the mtcars dataset
cylinder_counts <- table(data$cyl)  # Count the occurrences of each cylinder type
barplot(cylinder_counts, main = "Bar Plot of Cylinder Counts", xlab = "Number of Cylinders", ylab = "Frequency")

# 2. Box Plot (boxplot())
# Create a box plot for the mtcars dataset
boxplot(data, main = "Box Plot of mtcars Dataset", ylab = "Values")

# 3. Histogram (hist())
# Create a histogram for the 'mpg' column in the mtcars dataset
hist(data$mpg, main = "Histogram of MPG", xlab = "Miles per Gallon", col = "lightblue")

# 4. Scatter Plot Matrix (pairs())
# Create a scatter plot matrix for the first 4 columns of the mtcars dataset
pairs(data %>% select(1:4), main = "Scatter Plot Matrix of mtcars (First 4 Columns)")

# Custom Scatter Plot with Regression Line
# Create a scatter plot with a regression line for 'wt' vs 'mpg'
plot(data$wt, data$mpg, 
     main = "Scatterplot of Weight vs MPG with Regression Line",
     xlab = "Weight (1000 lbs)",
     ylab = "Miles per Gallon",
     pch = 19,  # Solid circle
     col = "blue")
abline(lm(mpg ~ wt, data = data), col = "red", lwd = 2)  # Add regression line

# Custom Panels in pairs()
# Define a custom histogram panel function
panel.hist <- function(x, ...) {
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

# Define a custom scatter plot panel function with regression line
panel.smooth <- function(x, y, ...) {
    points(x, y, ...)
    abline(lm(y ~ x), col = "blue")
}

# Create a scatter plot matrix with custom panels
pairs(data[1:5], 
      panel = panel.smooth,  # Custom scatter plot panel
      diag.panel = panel.hist,  # Custom histogram panel
      main = "Custom Scatter Plot Matrix with Histograms",
      pch = 24,  # Triangle
      bg = "light blue", 
      cex.labels = 2, 
      font.labels = 2)

```

## DataExplorer Package

```{r}
# Clear the environment
rm(list = ls())

# Load the DataExplorer package
# install.packages("pacman")
pacman::p_load(
  tidyverse,
  DataExplorer
)

# Load the mtcars dataset
d1 <- mtcars

# Display the first few rows of the dataset
head(d1)

# Plot structure of the dataset
DataExplorer::plot_str(d1)  # Visualize the structure of the dataset

# Display basic information about the dataset
DataExplorer::introduce(d1)  # Provide an overview of the dataset

# Plot correlation heatmap
DataExplorer::plot_correlation(d1)  # Visualize the correlation matrix as a heatmap

# Plot box plots grouped by 'disp' (displacement)
DataExplorer::plot_boxplot(d1, by = "disp")  # Create box plots for each variable grouped by displacement

# Plot bar charts
DataExplorer::plot_bar(d1)  # Create bar charts for categorical variables

# Plot density plots
DataExplorer::plot_density(d1)  # Create density plots for numerical variables

# Plot histograms
DataExplorer::plot_histogram(d1)  # Create histograms for numerical variables

# Plot missing values
DataExplorer::plot_missing(d1)  # Visualize missing values in the dataset

# Plot scatter plots grouped by 'cyl' (number of cylinders)
DataExplorer::plot_scatterplot(d1, by = "cyl")  # Create scatter plots for each variable grouped by the number of cylinders

# Plot QQ plots
DataExplorer::plot_qq(d1)  # Create QQ plots for numerical variables
```

## GGally Package：ggpairs()

Matrix Section

A pairwise matrix has three main sections:`lower`,`upper`,and`diag`.The`lower`and`upper`sections can contain three types of plots:`continuous`,`combo`,and`discrete`.The`diag`section can only contain`continuous`or`discrete`.

• continuous:Both X and Y are continuous variables.
• combo:One of X and Y is discrete,and the other is continuous.
• discrete:Both X and Y are discrete variables.
To adjust each section,you can supply a list of information.The list can contain the following elements:
• continuous:A string representing the tail end of a`ggally_NAME`function,or a custom function.
• combo:A string representing the tail end of a`ggally_NAME`function,or a custom function.(Not applicable for the`diag`list)
• discrete:A string representing the tail end of a`ggally_NAME`function,or a custom function.
• mapping:If mapping is provided,only the mapping for that section will be overwritten.
The current list of valid`ggally_NAME`functions can be seen in`vig_ggally("ggally_plots")`.

```{r message=FALSE}
# Load necessary libraries
# install.packages("pacman")
pacman::p_load(
  tidyverse,
  GGally
)

# Load the mtcars dataset
d1 <- mtcars

# Custom colors for the Iris dataset
# Create a pairwise plot matrix for the Iris dataset with custom colors
ggpairs(
  iris, 
  mapping = aes(color = Species),  # Color by species
  lower = list(continuous = wrap("smooth")),  # Lower triangle: smooth scatter plot
  progress = FALSE  # Disable progress bar
) + 
  scale_colour_manual(values = c('red', 'blue', 'orange')) +  # Custom colors for species
  theme_bw()  # Use a black and white theme

# Scatter plot matrix with regression lines, density plots, and correlation coefficients
# Create a scatter plot matrix for the first 4 columns of the mtcars dataset
ggpairs(
  mtcars %>% select(1:4),  # Select the first 4 columns
  title = "Pairs Plot of mtcars with Custom Panels",  # Title of the plot
  upper = list(continuous = wrap("cor", size = 5)),  # Upper triangle: correlation coefficients
  lower = list(continuous = wrap("smooth", alpha = 0.6), combo = "facetdensity"),  # Lower triangle: smooth scatter plot with density
  diag = list(continuous = wrap("densityDiag")),  # Diagonal: density plots
  progress = FALSE  # Disable progress bar
) + 
  theme_bw() +  # Use a black and white theme
  theme(plot.title = element_text(hjust = 0.5))  # Center the title

# Scatter plot matrix with regression lines, histograms, and correlation coefficients
# Create a scatter plot matrix for selected columns of the mtcars dataset
ggpairs(
  mtcars,
  columns = c("mpg", "wt", "hp", "drat"),  # Select specific columns
  lower = list(continuous = wrap("smooth", alpha = 0.6)),  # Lower triangle: smooth scatter plot
  diag = list(continuous = wrap("barDiag", binwidth = 1)),  # Diagonal: histograms
  progress = FALSE  # Disable progress bar
) +
  theme(plot.title = element_text(hjust = 0.5)) +  # Center the title
  theme_bw()  # Use a black and white theme

# Custom diagonal plotting function and create a scatter plot matrix
# Define a custom function for the diagonal panels
custom_diag <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_histogram(aes(y = ..density..), color = "black", ...) +  # Histogram with density
    geom_density(alpha = 0.5, ...)  # Overlay density plot
}

# Create a scatter plot matrix with custom diagonal panels
p <- ggpairs(
  d1,
  columns = c(1, 3:7),  # Select specific columns
  upper = list(continuous = wrap("cor", size = 8)),  # Upper triangle: correlation coefficients
  lower = list(continuous = wrap("smooth")),  # Lower triangle: smooth scatter plot
  diag = list(continuous = custom_diag),  # Diagonal: custom histogram and density plot
  progress = FALSE  # Disable progress bar
) %>% 
  print()  # Print the plot

```




























