---
title: "Correlation Analysis"
author: "Dean Hawk"
date: "2024-03-27"
output: html_document
---

# Correlation Analysis

## Preliminary Preparation

```{r}
# Load packages
pacman::p_load(ggcorrplot,corrmorant)
library(tidyverse)
library(ggplot2)
# install.packages("ggcorrplot")
library(ggcorrplot) # Heatmap
# install.packages("corrplot")
library(corrplot) # Heatmap
# install.packages("pheatmap")
library(pheatmap) # Heatmap
# install.packages("circlize")
library(circlize) # Chord diagram
# library(plotly,warn.conflicts = flase)
# devtools::install_github("thomasp85/scico")
library(scico) # Create a palette using the "scico" package
library(psych)
library(magrittr)
library(corrmorant) # Create visualizations of correlation matrices?

# Data loading
dt <- mtcars

# Convert the correlation matrix to a data frame
# correlation_df <- as.data.frame(correlation_matrix)
```

## Pearson Correlation CoefficientThe Pearson correlation coefficient is used to assess the strength and direction of the linear relationship between two continuous variables.

```{r}
# Calculate the Pearson correlation coefficient matrix using the cor function
pearson_cor_matrix <- cor(dt[, c("disp", "mpg", "cyl")],method = "pearson")
print(pearson_cor_matrix)
```

## Spearman Rank Correlation CoefficientThe Spearman rank correlation coefficient does not assume that the data are normally distributed. It is suitable for continuous or ordinal categorical variables and calculates correlation based on the ranks of the original data.

```{r}
# Calculate the Spearman rank correlation coefficient: spearman_cor <- cor(x, y, method = "spearman")
spearman_cor_matrix <-cor(dt[, c("disp", "mpg", "cyl")],method = "spearman")
print(spearman_cor_matrix)
```

## Kendall Rank Correlation CoefficientThe Kendall rank correlation coefficient is another non-parametric measure of correlation, suitable for small samples or ordinal categorical variables.

```{r}
# Calculate the Kendall rank correlation coefficient: kendall_cor <- cor(x, y, method = "kendall")
kendall_cor_matrix <- cor(dt[, c("disp", "mpg", "cyl")],method = "kendall")
print(kendall_cor_matrix)
```

## Visualizing Correlation Matrices (Correlation Heatmaps)If you have multiple variables and want to assess the correlations between them simultaneously, you can create a correlation matrix and visualize it using a heatmap.

```{r}
# Assume data is a data frame containing multiple variables
cor_matrix <- cor(dt, method = "pearson")

# Use the ggplot2 package to draw a heatmap
library(ggplot2)
library(reshape2)

# Convert the correlation matrix to long format
cor_melt <- melt(cor_matrix)

# Draw the heatmap
ggplot(cor_melt, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", limit = c(-1, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Use ggplot2 to draw a heatmap
ggplot(cor_melt, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "", fill = "Correlation")

# Draw a heatmap using ggcorrplot()
ggcorrplot(cor_matrix)
ggcorrplot(cor_matrix,method = "circle",lab=T)

# Draw a heatmap using corrplot()
corrplot(cor_matrix)

corrplot(cor_matrix, method = "circle", 
         tl.col = "black", tl.cex = 1.2, tl.srt = 45)

# Change parameters, method = "ellipse", type = "upper".
corrplot(cor_matrix, method = "ellipse", 
         type = "upper",
         tl.col = "black", tl.cex = 1.2, tl.srt = 45)
# method and type have other options, modify according to needs
# method = c("circle", "square", "ellipse", "number", "shade", "color", "pie")
# type = c("full", "lower", "upper")

# Combination, display correlation coefficients in the lower left and ellipse graphics in the upper right
corrplot(cor_matrix, method = "ellipse", type = "upper",
         tl.col = "black", tl.cex = 0.8, tl.srt = 45, tl.pos = "lt")

corrplot(cor_matrix, method = "number", type = "lower",
         tl.col = "n", tl.cex = 0.8, tl.pos = "n",add = T)

# Change color, add a custom color
addcol <- colorRampPalette(c("red", "white", "blue"))
corrplot(cor_matrix, method = "pie", type = "upper",col = addcol(100), 
         tl.col = "black", tl.cex = 0.8, tl.srt = 45,
         tl.pos = "lt")

corrplot(cor_matrix, method = "number", type = "lower",col = addcol(100), 
         tl.col = "n", tl.cex = 0.8, tl.pos = "n",
         add = T)

# Add statistical significance stars
testRes = cor.mtest(dt, method="pearson",conf.level = 0.95)
corrplot(cor_matrix, method = "color", col = addcol(100), 
         tl.col = "black", tl.cex = 0.8, tl.srt = 45,tl.pos = "lt",
         p.mat = testRes$p, diag = T, type = 'upper',
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 1.2,
         insig = 'label_sig', pch.col = 'grey20', order = 'AOE')

corrplot(cor_matrix, method = "number", type = "lower",col = addcol(100), 
         tl.col = "n", tl.cex = 0.8, tl.pos = "n",order = 'AOE',
         add = T)

```

```{r}
# Draw a heatmap using pheatmap()
scico_palette_show()
mycol <- scico(1000, palette = "vik") # Create a palette using the "scico" package

pheatmap(cor_matrix, 
         show_colnames = TRUE,   # Whether to display column names
         show_rownames=TRUE,     # Whether to display row names
         fontsize=13,             # Font size
         color = colorRampPalette(c('#00185A','#286391','#E9E1DB','#AB5A2B','#501008'))(50), # Specify the colors for the heatmap
         # color = mycol,
         annotation_legend=TRUE, # Whether to display the legend
         border_color=NA,        # Border color NA means no border
         scale="none",           # Specify the normalization method. "row" normalizes by row, "column" by column, "none" means no processing
         cluster_rows = TRUE,    # Whether to cluster rows
         cluster_cols = TRUE     # Whether to cluster columns
)

```

```{r}
# Convert the correlation coefficient matrix to wide format, with row names as environmental variables,
# column names as species, and values as correlation coefficients
rvalue <- df %>%
  select(1, 2, 3) %>%
  pivot_wider(names_from = genus, values_from = r) %>%
  column_to_rownames(var = "env")

# Convert the significance symbol matrix to wide format, with row names as environmental variables,
# column names as species, and values as significance symbols
pvalue <- df %>%
  select(1, 2, 5) %>%
  pivot_wider(names_from = genus, values_from = p_signif) %>%
  column_to_rownames(var = "env")

mycol <- scico(100, palette = "vik")

# Draw a heatmap, showing correlation coefficients, clustering rows and columns, no border,
# displaying p-values as numbers, setting font size and color for numbers,
# setting the main title to a space, setting cell width and height, and using a custom color map
pheatmap(cor_matrix, 
         scale = "none", 
         cluster_row = TRUE, 
         cluster_col = TRUE, 
         border = NA,
         # display_numbers = pvalue, 
         fontsize_number = 12, 
         number_color = "white",
         main = " ",
         cellwidth = 21, 
         cellheight = 20, 
         color = mycol
)
```













